---
title: "Glucose Project Version 6"
author: "Verena Steffen"
output: html_document
---

```{r,echo=FALSE}
# Clear workspace
rm(list=ls())
# Reset par
resetPar <- function() {
    dev.new()
    op <- par(no.readonly = TRUE)
    dev.off()
    op
}
par(resetPar())
# Set working directory
setwd("/home/vreni/Dropbox/Zürich/Uni Zürich Biostatistik Master/4_FS_2016/STA490_StatisticalConsulting")

# Load libraries
library("knitr")
library("lattice")
library("plyr")
library("dplyr")
library("lme4")
library("ggplot2")


# setwd("/home/c/vsteff/STA490_StatisticalConsulting/")
# Load data
# load("STA490/glu_v05.RData")

# Store data in list
# data <- list(glu = glu, diab = diab, diabetic.patients = diabetic.patients, npatients = npatients, outliers = outliers, percent.outliers = percent.outliers, total = total)
# 
# save(data, file = "glu_v06_list.RData")

load("STA490/glu_v07_list.RData")
glu <- data$glu
patients <- data$patients
timediff <- data$timediff
```

### Questions from Project Plan

*Anthropometrics*

- Age distribution
- Sex ratio

```{r}
# Age distribution
hist(glu$Age, main = "Age distribution")

# Sex ratio
sex.ratio <- sum(glu$Sex == "M") / sum(glu$Sex == "F")
plot(glu$Sex, main = "Sex ratio", sub = paste("ratio M/F =", round(sex.ratio,2)))

# Glucose per patient
ranges <- c("severe hypoglycemia", 
            "hypoglycemia",
            "euglycemia",
            "mild hyperglycemia",
            "pronounced hyperglycemia",
            "severe hyperglycemia")
low.borders <- c(2.2, 4, 7.7, 10, 15, range(glu$Glu)[2])
library(lattice)
p1 <- xyplot(Glu ~ Time, groups = PID, 
       data = glu[which(as.numeric(glu$PID) <= 50), ], 
       type = "b", pch = "*", col = rainbow(50), cex = 2,
       main = list("Glucose levels of 50 patients in 2014"), 
       xlab = list("Month 2014"),
       ylab = list("Glucose (mmol/l)"),
       scales = list(y = list(log = TRUE, equispaced.log = FALSE)),
       ylim = range(glu$Glu), 
       panel = function(...) {
           panel.xyplot(...)
           panel.abline(h = log(low.borders, 10), lty = "dotted", col = "black")
           panel.text(x = as.numeric(min(glu$Time)), 
                      y = log(0.9*low.borders, 10), 
                      ranges,  # text
                      pos = 4)}  # left-bound
       )
print(p1)
```


*Blood glucose control*

- Average BG (mmol/l)
- Average BG (mmol/l) based on A1C (Nathan et. al. Diabetes Care 2008)
- HbA1C (%)
- BG measurements/patient (n)
- % of values within glycemic ranges
- Intraindividual % of values within, below and above glycemic target (7.8-10 mmol/l)
- % of hypo- and hyperglycemic values
    - Stratified by age, gender, department
    - Diurnal variation (e.g. more hypoglycemic events at night?)
- Blood glucose variability: e.g. SD, CV

```{r}
# Average BG
summary(glu$Glu)

# Average BG based on HbA1c?

# HbA1c
summary(glu$HbA1c.DCCT)
summary(glu$HbA1c.IFCC)

# % of values within glycemic ranges
library(dplyr)
a <- glu %>% 
    group_by(glyc.control) %>% 
    summarize(n = n())
a
```



*Predictor of severe hypoglycaemia with BG<2.2 (e.g. prior BG≤3.9 during same admission)*

*Comparison between blood glucose control and comorbidity factors*

- Kidney dysfunction
- Liver dysfunction
- Dyslipidemia


### Models

```{r}
glu2 <- glu[which(as.numeric(glu$PID)<1000), ]
library(lme4)

# Fit model
mm1 <- lmer(log(Glu) ~ Lab + Sex + Ward + Age + (1 | PID), data = glu2, REML = FALSE)
plot(mm1, pch = "+")
par(mfrow=c(1,2))
acf(resid(mm1), pch = "+")
pacf(resid(mm1), pch = "+")
par(mfrow=c(1,1))

str(mm1)

# Compute heteroscedasticity-consistent covariance matrix
# library(sandwich)
# a <- vcovHC(mm1)
```


### Patient wise summary

```{r}
CV <- function(mean, sd) (sd/mean)*100

pat.mean <- tapply(glu$Glu, glu$PID, mean)  # glu mean of every patient
pat.cv <- tapply(glu$cv, glu$PID, mean)
pat.sex <- as.factor(tapply(as.integer(glu$Sex), glu$PID, max))
pat.sex <- factor(pat.sex, levels = c(1, 2), labels = c("F", "M"))
pat.age <- as.integer(tapply(glu$Age, glu$PID, mean))
pat.diab <- as.factor(tapply(glu$diabetic, glu$PID, mean))
pat.no.meas <- ddply(data.frame(glu$Inpatient, glu$PID),
                  ~ glu.PID, summarise,
                  no.meas = sum(as.integer(glu.Inpatient)-1))$no.meas

# Length of stay
stays <- aggregate(Glu ~ stay + PID, data = glu[which(glu$Inpatient == 1), ], 
               function(x) sd(x, na.rm=T)/mean(x,na.rm=T)*100)
hist(stays$Glu, xlim = c(0, 50), breaks = 30, xlab = "CV per stay")
head(stays, 20)

stays2 <- aggregate(Time ~ stay + PID, data = glu[which(glu$Inpatient == 1), ], 
               function(x) max(x) - min(x), units = "days")
head(stays2)




# timediff <- tapply(glu$Time, glu$PID, diff)
# timediff <- lapply(timediff, as.numeric, units = "days")
# head(timediff, 20)

patients <- data.frame(mean = pat.mean, 
                       cv = pat.cv,
                       sex = pat.sex,
                       age = pat.age,
                       no.meas = pat.no.meas,
                       diab = pat.diab)



```

```{r}
# Plots
nrow(patients)

hist(log(pat.mean))
hist(pat.age)
plot(pat.cv ~ pat.stay, pch = 16, cex = 0.5, col = "#EB811B98",
     main = "Patient CV vs. length of stay",
     xlab = "length of stay", ylab = "CV")
boxplot(pat.cv[which(pat.stay > 0)] ~ pat.stay[which(pat.stay > 0)], 
     pch = 16, cex = 0.5, col = "#EB811B98", 
     # log = "y",
     main = "Inpatient CV vs. length of stay",
     xlab = "length of stay (days)", ylab = "CV (%)")


library("ggplot2")
df <- data.frame(x = glu$Time, y = glu$Glu)
ggplot(df, aes(x=x,y=y)) + geom_point(shape=20,alpha=0.01) + theme_bw() + geom_smooth() + coord_trans(y = "log10") + ylim(5, 25)

ggplot(df,aes(x=x,y=y)) + stat_binhex() + theme_bw()

smoothScatter(df, log = "y", ylim = c(0.5, 50))
```



```{r}
# Store data in list
## data <- list(glu = glu, diab = diab, diabetic.patients = diabetic.patients, npatients = npatients, outliers = outliers, percent.outliers = percent.outliers, total = total, patients = patients)
# 
# glu$Ward <- gsub("\xd9", "ue", glu$Ward)
# glu$Ward <- as.factor(glu$Ward)
# data$glu <- glu
# save(data, file = "STA490/glu_v06_list.RData")
```


```{r}
# Flag public holidays
# glu$day <- weekdays(glu$Time)
# Mon <- which(glu$day == "Montag")
# Tue <- which(glu$day == "Dienstag")
# Wed <- which(glu$day == "Mittwoch")
# Thu <- which(glu$day == "Donnerstag")
# Fri <- which(glu$day == "Freitag")
# Sat <- which(glu$day == "Samstag")
# Sun <- which(glu$day == "Sonntag")
# days <- c(Mon = length(Mon),
#           Tue = length(Tue),
#           Wed = length(Wed),
#           Thu = length(Thu),
#           Fri = length(Fri),
#           Sat = length(Sat),
#           Sun = length(Sun))
# 
# public.holidays.bern2014 <- c("14-01-01",  # Neujahr
#                               "14-01-02",  # Berchtoldstag
#                               "14-04-18",  # Karfreitag
#                               "14-04-21",  # Ostermontag
#                               "14-05-29",  # Auffahrt
#                               "14-06-09",  # Pfingstmontag
#                               "14-08-01",  # Nationalfeiertag Schweiz
#                               "14-12-25",  # Weihnachten
#                               "14-12-26"   # Stephanstag
#                               )
# 
# l <- numeric(length(public.holidays.bern2014))
# ind.ph <- NULL
# for (i in (1:length(public.holidays.bern2014))) {
#   ind.ph <- c(ind.ph, which(startsWith(as.character(glu$Time),
#                                    public.holidays.bern2014[i])))
#   l[i] <- length(ind.ph)
# }
# 
# days <- c(days, "Publ. hol." = sum(l))
# barchart(days, xlab = list("Glucose measurements"), col = "grey")
# 
# glu$holiday <- ifelse(glu$day == "Sonntag", TRUE, FALSE)
# glu$holiday[ind.ph] <- TRUE
```

```{r}
# Define number of inpatient stays

# a <- Sys.time()
# stay <- list()
# for (i in 1:nrow(patients)) {
#   x <- glu[glu$PID == i, "Inpatient"]
#   y <- rle(as.integer(x)-1)
#   v <- NULL
#   for (j in (1:length(y$values))) {
#     v <- c(v, rep(y$values[j]*j, y$lengths[j]))
#     stay[[i]] <- v
#   }
# }
# glu$stay <- unlist(stay)
# b <- Sys.time()
# b-a
```




```{r}
# # Save Data
# data$timediff <- timediff
# data$glu <- glu
# save(data, file = "STA490/glu_v07_list.RData")
```

