---
title: "Glucose Project Version 4"
author: "Verena Steffen"
date: "April 5, 2016"
output: html_document
---
```{r,echo=FALSE}
# Clear workspace
rm(list=ls())
# Set working directory
setwd("/home/vreni/Dropbox/Zürich/Uni Zürich Biostatistik Master/4_FS_2016/STA490_StatisticalConsulting")
# Load data
load("STA490/old versions/glu_v03.RData")
```

# Exploratory plots
## Patient vs. time
```{r}
# Some plots
ranges <- c("severe hypoglycemia", 
            "hypoglycemia",
            "euglycemia",
            "mild hyperglycemia",
            "pronounced hyperglycemia",
            "severe hyperglycemia")
low.borders <- c(2.2, 4, 7.7, 10, 15, range(glu$Glu)[2])
library(lattice)
p1 <- xyplot(Glu ~ Time, groups = PID, 
       data = glu[which(as.numeric(glu$PID) <= 50), ], 
       type = "b", pch = "*", col = rainbow(50), cex = 2,
       main = list("Glucose levels of 50 patients in 2014", cex = 2), 
       xlab = list("Month 2014", cex = 1.3),
       ylab = list("Glucose (mmol/l)", cex = 1.3),
       scales = list(cex = 1.3, y = list(log = TRUE, equispaced.log = FALSE)),
       ylim = range(glu$Glu), 
       panel = function(...) {
           panel.xyplot(...)
           panel.abline(h = log(low.borders, 10), lty = "dotted", col = "black")
           panel.text(x = as.numeric(min(glu$Time)), 
                      y = log(0.9*low.borders, 10), 
                      ranges,  # text
                      pos = 4, 
                      cex = 1.3)}  # left-bound
       )
print(p1)
# png("STA490/plots/glucose_patient.png", width = 1300, height = 800); print(p1); dev.off()
```

## Other parameters vs. Glucose
Calculate correlations and summarize them.
```{r}
pars <- as.list(glu[5:17])
cors <- unlist(lapply(X = pars, FUN = cor, y = glu$Glu, use = "complete.obs"))

cors <- as.data.frame(cors)
cors$descr <- c("Creatinine POC (point of care)",
           "Creatinine CC (clin. chem.)",
           "estimated Glomerular Filtration Rate, MDR formula",
           "estimated Glomerular Filtration Rate, CKDE formula", 
           "Hemoglobin A1c",
           "Hemoglobin A1c", 
           "aspartate amino transferase",
           "alanine amino transferase",
           "total cholesterol",
           "Triglygerides",
           "high-density lipoprotein",
           "low-density lipoprotein",
           "low-density lipoprotein, calculated")
cors$func <- c(rep("Kidney", 4), 
          rep("Diabetic Control", 2), 
          rep("Liver", 2),
          rep("Fat", 5))
colnames(cors) <- c("Pearson Cor. w/ Glucose", 
                    "Parameter",
                    "Met. Function")
knitr::kable(cors)
```

### Plot some eventually relevant parameters.
#### HbA1c.DCCT: Hemoglobin A1c  /// Diabetic Control
```{r}
xyplot(HbA1c.DCCT ~ Glu, data = glu, pch = "*", cex = 2, 
       scales = list(list(log = TRUE, equispaced.log = FALSE)))
```

#### HbA1c.IFCC: Hemoglobin A1c  /// Diabetic Control (use that!)
```{r}
xyplot(HbA1c.IFCC ~ Glu, data = glu, pch = "*", cex = 2, 
       scales = list(list(log = TRUE, equispaced.log = FALSE)))
```

#### CHOL: total cholesterol  /// Fat
```{r}
xyplot(CHOL ~ Glu, data = glu, pch = "*", cex = 2, 
       scales = list(list(log = TRUE, equispaced.log = FALSE)))
```

#### TG: Triglygerides  /// Fat
```{r}
xyplot(TG ~ Glu, data = glu, pch = "*", cex = 2, 
       scales = list(list(log = TRUE, equispaced.log = FALSE)))
```

#### HDL: high-density lipoprotein  /// Fat
```{r}
xyplot(HDL ~ Glu, data = glu, pch = "*", cex = 2, 
       scales = list(list(log = TRUE, equispaced.log = FALSE)))
```

#### LDL: low-density lipoprotein  /// Fat
```{r}
xyplot(LDL ~ Glu, data = glu, pch = "*", cex = 2, 
       scales = list(list(log = TRUE, equispaced.log = FALSE)))
```

#### LDL: low-density lipoprotein, calculated  /// Fat
```{r}
xyplot(LDL.calc ~ Glu, data = glu, pch = "*", cex = 2, 
       scales = list(list(log = TRUE, equispaced.log = FALSE)))
```


## Distribution of Glucose
```{r}
# Glucose distribution
hist(glu$Glu)

# Log transformation
hist(log(glu$Glu), prob = TRUE, ylim = c(0,1.2))
x<-seq(0,1000,0.01)
curve(dnorm(x, mean=mean(log(glu$Glu), na.rm=T), sd=sd(log(glu$Glu), na.rm=T)), col = 3, add = TRUE)
lines(density(log(glu$Glu), na.rm=T, from = min(log(glu$Glu), na.rm=T), to = max(log(glu$Glu), na.rm=T)), col = 2)
legend("topleft", c("Normal Density", "True Density"), lty = 1, col = c(3, 2), box.lty = 0)

qqnorm(log(glu$Glu), pch = ".", cex = 3)
qqline(log(glu$Glu), col = 2)

# Sqrt transformation
hist(sqrt(glu$Glu), prob = TRUE, ylim = c(0,1.2))
x<-seq(0,1000,0.01)
curve(dnorm(x, mean=mean(sqrt(glu$Glu), na.rm=T), sd=sd(sqrt(glu$Glu), na.rm=T)), col = 3, add = TRUE)
lines(density(sqrt(glu$Glu), na.rm=T, from = min(sqrt(glu$Glu), na.rm=T), to = max(sqrt(glu$Glu), na.rm=T)), col = 2)
legend("topright", c("Normal Density", "True Density"), lty = 1, col = c(3, 2), box.lty = 0)

qqnorm(sqrt(glu$Glu), pch = ".", cex = 3)
qqline(sqrt(glu$Glu), col = 2)
```

--> Under a log transformation, a linear mixed model should be suitable.


## Distribution of Age      
- Assumption: none of the patients is > 100 years old.
- patients < 3 years are excluded from the analysis.

```{r}
hist(glu$Age)
```





