---
title: "Glucose Project Version 6"
author: "Verena Steffen"
output: html_document
---

```{r,echo=FALSE}
# Clear workspace
rm(list=ls())
# Reset par
resetPar <- function() {
    dev.new()
    op <- par(no.readonly = TRUE)
    dev.off()
    op
}
par(resetPar())
# Set working directory
setwd("/home/vreni/Dropbox/Zürich/Uni Zürich Biostatistik Master/4_FS_2016/STA490_StatisticalConsulting")
load("STA490/glu_v04.RData")
```

### Questions from Project Plan

*Anthropometrics*

- Age distribution
- Sex ratio

```{r}
# Age distribution
hist(glu$Age, main = "Age distribution")

# Sex ratio
sex.ratio <- sum(glu$Sex == "M") / sum(glu$Sex == "F")
plot(glu$Sex, main = "Sex ratio", sub = paste("ratio M/F =", round(sex.ratio,2)))

# Glucose per patient
ranges <- c("severe hypoglycemia", 
            "hypoglycemia",
            "euglycemia",
            "mild hyperglycemia",
            "pronounced hyperglycemia",
            "severe hyperglycemia")
low.borders <- c(2.2, 4, 7.7, 10, 15, range(glu$Glu)[2])
library(lattice)
p1 <- xyplot(Glu ~ Time, groups = PID, 
       data = glu[which(as.numeric(glu$PID) <= 50), ], 
       type = "b", pch = "*", col = rainbow(50), cex = 2,
       main = list("Glucose levels of 50 patients in 2014"), 
       xlab = list("Month 2014"),
       ylab = list("Glucose (mmol/l)"),
       scales = list(y = list(log = TRUE, equispaced.log = FALSE)),
       ylim = range(glu$Glu), 
       panel = function(...) {
           panel.xyplot(...)
           panel.abline(h = log(low.borders, 10), lty = "dotted", col = "black")
           panel.text(x = as.numeric(min(glu$Time)), 
                      y = log(0.9*low.borders, 10), 
                      ranges,  # text
                      pos = 4)}  # left-bound
       )
print(p1)
```


*Blood glucose control*

- Average BG (mmol/l)
- Average BG (mmol/l) based on A1C (Nathan et. al. Diabetes Care 2008)
- HbA1C (%)
- BG measurements/patient (n)
- % of values within glycemic ranges
- Intraindividual % of values within, below and above glycemic target (7.8-10 mmol/l)
- % of hypo- and hyperglycemic values
    - Stratified by age, gender, department
    - Diurnal variation (e.g. more hypoglycemic events at night?)
- Blood glucose variability: e.g. SD, CV

```{r}
# Average BG
summary(glu$Glu)

# Average BG based on HbA1c?

# HbA1c
summary(glu$HbA1c.DCCT)
summary(glu$HbA1c.IFCC)

# % of values within glycemic ranges
library(dplyr)
a <- glu %>% 
    group_by(glyc.control) %>% 
    summarize(n = n())
a

```



*Predictor of severe hypoglycaemia with BG<2.2 (e.g. prior BG≤3.9 during same admission)*

*Comparison between blood glucose control and comorbidity factors*

- Kidney dysfunction
- Liver dysfunction
- Dyslipidemia


### Models

```{r}
glu2 <- glu[which(as.numeric(glu$PID)<1000), ]
library(lme4)

# Fit model
mm1 <- lmer(log(Glu) ~ Lab + Sex + Ward + Age + Time.trunc + (1 | PID), data = glu2, REML = FALSE)
plot(mm1, pch = "+")
par(mfrow=c(1,2))
acf(resid(mm1), pch = "+")
pacf(resid(mm1), pch = "+")
par(mfrow=c(1,1))

# Compute heteroskedasticity-consistent covariance matrix
# library(sandwich)
# a <- vcovHC(mm1)


```






